---
title: "advancedgraphingtwo"
output: html_document
date: "2024-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Okay, let's make some fake data!


```{r}
library(tidyverse)
library(lubridate)
library(magrittr)
```

```{r}
set.seed(123) # for reproducibility
```

```{r}
# Generate fake data
shopping_data <- tibble(
  Transaction = 1:2000,
  Purchase_Date = sample(seq(as.Date('2020-01-01'), as.Date('2023-12-31'), by="day"), 2000, replace = TRUE),
  Customer_ID = 1000 + seq(1:2000),
  Gender = sample(c("M", "F"), 2000, replace = TRUE),
  Marital_Status = sample(c("S, M"), 2000, replace = TRUE),
  Homeowner = sample(c("Y", "N"), 2000, replace = TRUE),
  Children = sample(c("Y", "N"), 2000, replace = TRUE),
  Annual_Income = sample(c("$30,000 - $50,000", "$50,000 - $70,000", "$70,000 - $90,000", "$90,000 - $110,000"), 2000, replace = TRUE),
  City = sample(c("New York", "Los Angeles", "Chicago", "Houston", "Phoenix", "Philadelphia", 
                  "San Antonio", "San Diego", "Dallas", "San Jose", "Austin", "Jacksonville", 
                  "San Francisco", "Indianapolis", "Columbus", "Fort Worth", "Charlotte", 
                  "Seattle", "Denver", "El Paso", "Detroit", "Washington", "Boston"), 2000, replace = TRUE),
  State_or_Province = sample(c("NY", "CA", "IL", "TX", "AZ", "PA", "TX", "CA", "TX", "CA", "TX", "FL", 
                               "CA", "IN", "OH", "TX", "NC", "WA", "CO", "TX", "MI", "DC", "MA"), 2000, replace = TRUE),
  Country = "USA",
  Product_Family = sample(rep(c("Electronics", "Clothing", "Books", "Furniture", "Appliances", "Toys"), each = 333), 2000, replace = TRUE),
  Product_Department = case_when(
    Product_Family == "Electronics" ~ sample(c("TVs", "Laptops"), 2000, replace = TRUE),
    Product_Family == "Clothing" ~ sample(c("Shirts", "Pants"), 2000, replace = TRUE),
    Product_Family == "Books" ~ sample(c("Fiction", "Non-fiction"), 2000, replace = TRUE),
    Product_Family == "Furniture" ~ sample(c("Sofas", "Tables"), 2000, replace = TRUE),
    Product_Family == "Appliances" ~ sample(c("Refrigerators", "Microwaves"), 2000, replace = TRUE),
    Product_Family == "Toys" ~ sample(c("Dolls", "Action Figures"), 2000, replace = TRUE)
  ),
  Product_Category = sample(c("High-end", "Mid-range", "Low-end"), 2000, replace = TRUE),
  Units_Sold = sample(1:10, 2000, replace = TRUE),
  Revenue = ifelse(City %in% c("New York", "Los Angeles", "Chicago", "San Francisco", "Washington"), 
                   rnorm(2000, mean = 150, sd = 20), 
                   ifelse(City %in% c("Houston", "Phoenix", "Philadelphia", "San Antonio", "San Diego"), 
                          rnorm(2000, mean = 100, sd = 20), 
                          ifelse(City %in% c("Detroit", "Boston", "Jacksonville", "El Paso"), 
                          rnorm(2000, mean = 30, sd = 20),
                                rnorm(2000, mean = 60, sd = 20))))
)

```

Create a new df with total revenue for each city.  Change the data type of the City column to factor.
```{r}
city_rev <- shopping_data %>%
        group_by(City) %>%
        summarise(Revenue = sum(Revenue, na.rm = TRUE)) %>%
        arrange(Revenue) %>%
        mutate(City = factor(City, levels = .$City))
```


```{r}
ggplot(city_rev, aes(City, Revenue)) +
        geom_bar(stat = "identity") +
        coord_flip()
```
If you provide the argument stat=”identity” to geom_bar() then you’re telling R to calculate the sum of the y variable, grouped by the x variable and use bars to display the sums.


```{r}
ggplot(city_rev, aes(Revenue, City)) +
        geom_point()
```

Create anothe df with total revenue for each gender in eaxh city. Change the data type of the City column to factor again.
```{r}
city_gender_rev <- shopping_data %>%
        group_by(City, Gender) %>%
        summarise(Revenue = sum(Revenue, na.rm = TRUE)) %>%
        ungroup() %>%
        mutate(City = factor(City, levels = city_rev$City))
```


```{r}
ggplot(city_gender_rev, aes(City, Revenue, fill = Gender)) +
        geom_bar(stat = "identity")+
        coord_flip()

```

```{r}
ggplot(city_gender_rev, aes(City, Revenue, fill = Gender)) +
        geom_bar(stat = "identity", position = "dodge") +
        coord_flip()
```

```{r}
ggplot(city_gender_rev, aes(City, Revenue, fill = Gender)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        facet_wrap(~ Gender)
```


```{r}
ggplot(city_gender_rev, aes(Revenue, City)) +
        geom_point(aes(color = Gender))
```

```{r}
ggplot(city_gender_rev, aes(Revenue, City)) +
        geom_line(aes(group = City)) +
        geom_point(aes(color = Gender))
```
```{r}
ggplot(city_gender_rev, aes(Revenue, City, label = round(Revenue, 0))) +
        geom_line(aes(group = City)) +
        geom_point(aes(color = Gender)) +
        geom_text(aes(color = Gender), size = 3)
```


Here we are creating two new dfs just for labels.
```{r}
right_label <- city_gender_rev %>%
        group_by(City) %>%
        arrange(desc(Revenue)) %>%
        top_n(1)   #Selects the first value for each city, which is he higher revenue

left_label <- city_gender_rev %>%
        group_by(City) %>%
        arrange(desc(Revenue)) %>%
        slice(2)   #Selects the second value for each city, which is the lower revenue

ggplot(city_gender_rev, aes(Revenue, City)) +
        geom_line(aes(group = City)) +
        geom_point(aes(color = Gender), size = 1.5) +
        geom_text(data = right_label, aes(color = Gender, label = round(Revenue, 0)),
                  size = 3, hjust = -.5) +
        geom_text(data = left_label, aes(color = Gender, label = round(Revenue, 0)),
                  size = 3, hjust = 1.5) +
        scale_x_continuous(limits = c(-500, 10500))
```


This code is finding cities where there is a significant difference in revenue between genders, by comparing the maximum and minimum revenue for each city and filtering out those where the difference is less than 20%.

Then we are using that new df to edit the left_label and right_label dfs and add some highlighting. 

```{r}
twenty_diff <- city_gender_rev %>% 
        spread(Gender, Revenue) %>% 
        group_by(City) %>% 
        mutate(Max = max(F, M),
               Min = min(F, M),
               Diff = Max / Min - 1) %>% 
        arrange(desc(Diff)) %>%
        filter(Diff > .2)
```
Explanation of the first section of that code:

1. `spread(Gender, Revenue)`: This function spreads the data frame from long to wide format, where each unique value of `Gender` becomes a new column, and its corresponding `Revenue` value populates that column. 
2. `group_by(City)`: This function groups the data by the `City` column. This means that all rows with the same city name will be grouped together.

3. `mutate(Max = max(F, M), Min = min(F, M), Diff = Max / Min - 1)`: This line creates three new columns: `Max`, `Min`, and `Diff`. 
   - `Max = max(F, M)` finds the maximum revenue value among the "Female" (`F`) and "Male" (`M`) columns for each city. 
   - `Min = min(F, M)` finds the minimum revenue value among the "Female" and "Male" columns for each city. 
   - `Diff = Max / Min - 1` calculates the difference between the maximum and minimum revenue, normalized by the minimum revenue, and subtracts 1. This gives us the percentage difference between the maximum and minimum revenue.

4. `arrange(desc(Diff))`: This function arranges the rows in descending order based on the `Diff` column, which represents the percentage difference between the maximum and minimum revenue. This means that cities with the largest differences in revenue will appear first.

5. `filter(Diff > .2)`: This function filters the rows to keep only those where the `Diff` column (the percentage difference between maximum and minimum revenue) is greater than 0.2 (20%). This filters out cities where the difference in revenue between genders is less than 20%.


```{r}
# filter the label data frames to only include those cities where the difference exceeds 20%
right_label <- filter(right_label, City %in% twenty_diff$City)
left_label <- filter(left_label, City %in% twenty_diff$City)

# filter the main data frame to only include those cities where the 
# difference exceeds 20%.
highlight <- filter(city_gender_rev, City %in% twenty_diff$City)


#plot the data using alpha of .3 to make everything 30% opaque, then plot highlight data over that at 100% opacity
ggplot(city_gender_rev, aes(Revenue, City)) +
        geom_line(aes(group = City), alpha = .3) +
        geom_point(aes(color = Gender), size = 1.5, alpha = .3) +
        geom_line(data = highlight, aes(group = City)) +
        geom_point(data = highlight, aes(color = Gender), size = 2) +
        geom_text(data = right_label, aes(color = Gender, label = round(Revenue, 0)),
                  size = 3, hjust = -.5) +
        geom_text(data = left_label, aes(color = Gender, label = round(Revenue, 0)),
                  size = 3, hjust = 1.5) +
        scale_x_continuous(limits = c(-500, 10500))
```


```{r}
# create a new label data frame
plot_label <- twenty_diff %>%
        select(City, Revenue = Max, Diff) %>%
        right_join(right_label)

p <- ggplot(city_gender_rev, aes(Revenue, City)) +
        geom_line(aes(group = City), alpha = .3) +
        geom_point(aes(color = Gender), size = 1.5, alpha = .3) +
        geom_line(data = highlight, aes(group = City)) +
        geom_point(data = highlight, aes(color = Gender), size = 2) +
        geom_text(data = plot_label, aes(color = Gender, 
                                         label = paste0("+", scales::percent(round(Diff, 2)))),
                  size = 3, hjust = -.5)

p

```

```{r, fig.height=10, fig.width=6}

p + scale_color_discrete(labels = c("Female", "Male")) +
        scale_x_continuous(labels = scales::dollar, expand = c(0.02, 0), 
                           limits = c(0, 8000),
                           breaks = seq(0, 8000, by = 2000)) +
        scale_y_discrete(expand = c(.02, 0)) +
        labs(title = "Total Revenue by City and Gender",
             subtitle = "Out of 23 cities, eleven locations experience a 20% or greater difference \nin revenue generated by males versus females. Chicago experiences the \ngreatest difference with females generating 86% more revenue than males.") +
        theme_minimal() +
        theme(axis.title = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank(),
              legend.title = element_blank(),
              legend.justification = c(0, 1), 
              legend.position = c(.1, 1.075),
              legend.background = element_blank(),
              legend.direction="horizontal",
              text = element_text(family = "Cambria"),
              plot.title = element_text(size = 20, margin = margin(b = 10)),
              plot.subtitle = element_text(size = 10, color = "darkslategrey", margin = margin(b = 25)),
              plot.caption = element_text(size = 8, margin = margin(t = 10), color = "grey70", hjust = 0))

```

```{r}
library(extrafont)
font_import()
```
```{r}
loadfonts(device = "win")
```