---
title: "advancedgraphingone"
output: html_document
date: "2024-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Our main packages
```{r}
library(dplyr)          # for data manipulation
library(tidyr)          # for data tidying
library(ggplot2)        # for generating the visualizations
```


Some basics:
```{r}
# create canvas
ggplot(mpg)

# variables of interest mapped
ggplot(mpg, aes(x = displ, y = hwy))

# data plotted
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point()
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(color = "blue")

```

The most common distinction between plots is what geometric objects (geoms) they include. ggplot2 supports a number of different types of geoms, including:

    geom_point for drawing individual points (e.g., a scatter plot)
    geom_line for drawing lines (e.g., for a line charts)
    geom_smooth for drawing smoothed lines (e.g., for simple trends or approximations)
    geom_bar for drawing bars (e.g., for bar charts)
    geom_histogram for drawing binned values (e.g. a histogram)
    geom_polygon for drawing arbitrary shapes
    geom_map for drawing polygons in the shape of a map! (You can access the data to use for these maps by using the map_data() function).

Each geom has some unique aesthetic properties.  For example, geom_line has the line_type property.
```{r}
#x and y mapping needed!
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth()

#no y mapping needed!
ggplot(data = mpg, aes(x = class)) +
  geom_bar()  

ggplot(data = mpg, aes(x = hwy)) +
  geom_histogram() 
```



Data we'll be using (Included in ggplot)
```{r}
head(midwest)
```

Filtering the top 25 Ohio counties in terms of college education
```{r}
ohio_top25 <- midwest %>%
        filter(state == "OH") %>%               #filter for OH
        select(county, percollege) %>%          #select these two columns
        arrange(desc(percollege)) %>%           #arrange them in descending order
        top_n(25) %>%                           #take the top 25
        arrange(percollege) %>%                 #arrange them in ascending order (default is ascending)
        mutate(county = factor(county, levels = .$county)) #change the county from a chr to factor
```


```{r}
ggplot(ohio_top25, aes(county, percollege)) +
        geom_bar(stat = "identity") +
        coord_flip()

ggplot(ohio_top25, aes(percollege, county)) +
        geom_point()
```
```{r}
ggplot(ohio_top25, aes(percollege, county)) +
        geom_segment(aes(x = 0, y = county, xend = percollege, yend = county), color = "grey50") +
        geom_point()
```
Now we're looking at the same percollege data for Ohio, but we're adding a mean column and and Aobve column to reference the mean for each county.
```{r}
ohio <- midwest %>%
        filter(state == "OH") %>%
        select(county, percollege) %>%
        arrange(percollege) %>%
        mutate(Avg = mean(percollege, na.rm = TRUE),
               Above = ifelse(percollege - Avg > 0, TRUE, FALSE),
               county = factor(county, levels = .$county))
```

```{r}
ggplot(ohio, aes(percollege, county, color = Above)) +
        geom_segment(aes(x = Avg, y = county, xend = percollege, yend = county), color = "grey50") +
        geom_point()
```

```{r}
top10 <- midwest %>%
        select(state, county, percollege) %>%
        group_by(state) %>%
        arrange(desc(percollege)) %>%
        top_n(10) %>%
        arrange(percollege) %>%
        unite(county_st, county, state, remove = FALSE) %>%
        mutate(county_st = factor(county_st, levels = .$county_st))
```

```{r}
ggplot(top10, aes(percollege, county_st)) +
         geom_segment(aes(x = 0, y = county_st, xend = percollege, yend = county_st), color = "grey50") +
        geom_point() +
        scale_y_discrete(labels = abbreviate) +
        facet_wrap(~ state, scales = "free_y")
```

```{r}
OH_top10 <- midwest %>%
        select(state, county, percollege) %>%
        filter(state == "OH") %>%
        arrange(desc(percollege)) %>%
        top_n(10) %>%
        arrange(percollege) %>%
        mutate(county = factor(county, levels = .$county))
```

```{r}
ggplot(OH_top10, aes(percollege, county, label = paste0(round(percollege, 0), "%"))) +
         geom_segment(aes(x = 0, y = county, xend = percollege, yend = county), color = "grey50") +
        geom_point(size = 7) +
        geom_text(color = "white", size = 2)
```

```{r}
ggplot(ohio, aes(percollege/100, county, color = Above)) +
        geom_segment(aes(x = Avg/100, y = county, xend = percollege/100, yend = county), color = "grey50") +
        geom_point() +
        annotate("text", x = .25, y = "ALLEN", label = "Above Average", color = "#00BFC4", size = 3, hjust = -0.1, vjust = .75) +
        annotate("text", x = .25, y = "FULTON", label = "Below Average", color = "#F8766D", size = 3, hjust = -0.1, vjust = -.1) +
        geom_segment(aes(x = .25, xend = .25 , y = "ASHLAND", yend = "DEFIANCE"),
                     arrow = arrow(length = unit(0.2,"cm")), color = "#00BFC4") +
        geom_segment(aes(x = .25, xend = .25 , y = "KNOX", yend = "PUTNAM"),
                     arrow = arrow(length = unit(0.2,"cm")), color = "#F8766D") +
        scale_x_continuous(labels = scales::percent, expand = c(0, 0), limits = c(.07, .33)) +
        labs(title = "College Educated Adults in Ohio Counties",
             subtitle = "The average percent of college educated adults in Ohio is 16.89%. Franklin, Greene, Geauga, and \nDelaware counties lead Ohio with over 30% of their adults being college educated while Vinton, \nAdams, Holmes, and Perry trailing with less than 10% of their adults being college educated.",
             caption = "U.S. Census Bureau: 2000 Census") +
        theme_minimal() +
        theme(axis.title = element_blank(),
              panel.grid.minor = element_blank(),
              legend.position = "none",
              text = element_text(family = "Georgia"),
              axis.text.y = element_text(size = 8),
              plot.title = element_text(size = 20, margin = margin(b = 10), hjust = 0),
              plot.subtitle = element_text(size = 12, color = "darkslategrey", margin = margin(b = 25, l = -25)),
              plot.caption = element_text(size = 8, margin = margin(t = 10), color = "grey70", hjust = 0))
```

